Should I Airbnb My Property: Analytic Code Explanation

### 2018-05-21 scripts/4_data_analysis.Rmd
Re-org Analysis, move viz, etc.

---
title: "Should I Airbnb My Property: Analytic Code Explanation"
author: "Andy Krause and Gideon Aschwanden"
date: "March 10, 2017"
output: html_document
---

@@

We load the workspace from the data clean script that includes the prepared data objects.

```{r load_str_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Property Information

  load(file.path(data.path, 'prepared', paste0('prep_workspace_', data_name, '.RData')))

@@

  str_df$cpppd <- 4.5
  str_df$avg.guests <- str_df$bedrooms * 1.5

```{r obs_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_df$str.obs.revenue <- ((str_df$bookings * str_df$med.rate) - 
                                 (str_df$bookings * str_df$avg.guests *
                                    str_df$cpppd))

@@

  str_df$extr.par <- 366 / str_df$total.days

  str_df$str.act.revenue <- ((str_df$bookings * str_df$extr.par *
                                  str_df$med.rate) - 
                                 (str_df$bookings * str_df$avg.guests * 
                                    str_df$cpppd * str_df$extr.par))
```

@@

  str_df$occ.rate <- 1 - (str_df$block.rate + str_df$avail.rate)

  str_df$pot.occ.rate <- ((str_df$bookings * str_df$extr.par) / 
    (366 - (366 * str_df$block.rate)))

  str_df$str.pot.revenue <- ((str_df$pot.occ.rate * 366 * str_df$med.rate) - 
                                 (str_df$pot.occ.rate * 366 * 
                                    str_df$avg.guests * str_df$cpppd))

  ltr.data$ltr.revenue <- ltr.data$gross.revenue

@@

@@

  ltr.mod.spec <- formula(log(event.price) ~ as.factor(type) + 
                            as.factor(bedbath) + as.factor(suburb) + 
                            as.factor(ltr.month))

@@

  imp.data <- imputeLtrRents(ltr.df=ltr.data,
                               str.df=str_df,
                               mod.spec=ltr.mod.spec,
                               match.factor='suburb')

@@

  # Extract out DFs (only those that imputed)
  str.metro <- merge(str_df, imp.data$imp.rent, by='property.id')

Next we impute the likely long-term search cost (days on market) for the short-term properties.

  # Apply dom to abb
  str.metro$imp.dom <- imputeDOM(str.metro, ltr.data, calc.type='median')

With the imputed DOMs and weekly rental rates we now make the an estimate of likely annual revenue if leased on the long term market. We append this to the original dataset.

  str.metro$ltr.imp.revenue <- str.metro$imp.rent * (52 - str.metro$imp.dom / 7)

Finally we make a comparison of the observed short term revenues and the imputed long term revenue and add these field back onto the original data.

  comp.revs <- compareRevenues(str.metro)

  str.metro <- merge(str.metro, comp.revs, by='property.id')

The above process of long-term revenue estimation treated the entire market as operating similarly. Realizing that the imputed long term rental values and search times may be more accurate when estimated at smaller scales, we functionalize the above processes and then apply it to short-term rentals, broken into submarkets by type, bed/bath and geographic area.  In other words, we separately estimate LTR rents for Apartments with 1bed/1bath in the city core, Apartments with 1bed/1bath in the beach submarket, etc. for all compbinations of type, bed/bath and geographic area. We then combine the results form all estimates back into a single data.frame that contains all of the short-term rentals. We use these finer scale estimates going forward (the data in the **str_df.ic** object).

```{r build.tbg.data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

 # Create the type, bed/bath and geo sub market variable

 str_df$tbg <- paste0(str_df$product,'.', str_df$geo.mrkt)
 ltr.data$tbg <- paste0(ltr.data$product,'.', ltr.data$geo.mrkt)

 # Make IC estimates at TBG level

 str_df.ic <- abbImputeCompare(str.df=str_df,
                                 ltr.df=ltr.data,
                                 mod.spec=ltr.mod.spec,
                                 match.factor='suburb',
                                 split.field='tbg',
                                 verbose=FALSE)

Having calculated the revenues, both short-term and long-term imputed, we now compare the actual (extr), potential and imputed long-term revenue distributions for the short term properties

```{r plot_revdens, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=6,fig.height=3.5}

  # Actual
   stra.rev <- str_df.ic[,c('property.id', 'str.act.revenue')]
   stra.rev$tenure <- 'Short-Term (Actual)   '
   names(stra.rev)[2] <- 'revenue'

  # Potential
   strp.rev <- str_df.ic[,c('property.id', 'str.pot.revenue')]
   strp.rev$tenure <- 'Short-Term (Potential)   '
   names(strp.rev)[2] <- 'revenue'

  # Long-Term
   ltr.rev <- str_df.ic[,c('property.id', 'ltr.imp.revenue')]
   ltr.rev$tenure <- 'Long-Term   '
   names(ltr.rev)[2] <- 'revenue'

  # Combine
   revdens.data <- rbind(stra.rev, strp.rev, ltr.rev)

  ## Build a plot

   rd.plot <- ggplot(revdens.data, 
                     aes(x=revenue, fill=tenure, color=tenure)) +
     geom_density(alpha=.5) +
     scale_fill_manual(values=c(ltr.col, abb.col[2], str.col)) +
     scale_color_manual(values=c(ltr.col, abb.col[2], str.col)) +
     xlab('\nAnnual Revenue') +
     scale_x_continuous(breaks=c(seq(0, 75000, by=25000)),
                        labels=c('$0', '$25k', '$50k', '$75k'))+
     theme(legend.position='none',
           legend.title = element_blank(),
           plot.title = element_text(hjust = 0.5),
           axis.text.y = element_blank(),
           axis.ticks.y = element_blank(),
           axis.title.y=element_blank()) +
     coord_cartesian(xlim=c(0, 85000))
     rd.plot <- rd.plot + facet_wrap(~tenure)

  ## Plot

     rd.plot

## Results

We will begin by analyzing the above estimates at the entire metro scale. Analyses of various submarkets will follow. Overall, we see that 31.1\% of properties raised more revenue via short term rental than expected in the long term market.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

  table(str_df.ic$str.act.pref) / nrow(str_df.ic)

Next, we plot the \% of properties that have short-term preference versus the occupancy rate as occupancy rate is an important determinant of short-term preference.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE}

 ## Set Preference data field

  str_df.ic$pref <- str_df.ic$str.act.pref

 ## Calculate the percentile fields

  str_df.ic$occ.qtl <- makeWtdQtl(str_df.ic$occ.rate, 
                                      return.type='rank') 
  str_df.ic$rate.qtl <- makeWtdQtl(str_df.ic$med.rate, 
                                  return.type='rank')

 ## Calculate raw and percentile occupancies

  rawocc.metro <- abbPrefPlot(str_df.ic,
                         x.field='occ.rate',
                         smooth=TRUE,
                         smooth.span=.5)

  qtlocc.metro <- abbPrefPlot(str_df.ic,
                        x.field='occ.qtl',
                        smooth=TRUE,
                        smooth.span=.5)

   ## Make plots

  # Raw
  raw.occ.metro <- ggplot(rawocc.metro$data,
                   aes(x=x.var, y=Var))+
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none') + 
    ggtitle('Raw Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

  # Quantile
  qtl.occ.metro <- ggplot(qtlocc.metro$data,
                   aes(x=x.var, y=Var))+
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Pcntl)') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none') + 
    ggtitle('Percentile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

  # Make Plot
  ggMultiPlots(raw.occ.metro, qtl.occ.metro, cols=2)

Next we build a heat map that shows which combinations of occupancy rate and nightly rate are most likely to result in short-term preference.

  rate.hm <- abbHeatMap(str_df.ic,
                x.field='occ.rate',
                y.field='med.rate',
                pref.field='str.act.pref',
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))
  rate.hm + ylab('Nightly Rate')

To make the distinction more clear, we employ a Support Vector Machine to classify the areas into short and long term preference.

  rate.hm.svm <- abbHeatMap(str_df.ic,
                            x.field='occ.rate',
                            y.field='med.rate',
                            pref.field='str.act.pref',
                            svm=T, 
                            alpha.count=F,
                            add.points=T,
                            fill.colors=c(abb.col[1], abb.col[5]))
  rate.hm.svm+ ylab('Nightly Rate')

As the distributions of occupancy rate and nightly rate are not uniform, we standardize these values to a percentile and then replot to give a better sense of the relative level of occupancy rate and nightly rate required for short-term preference.

  qtl.hm <- abbHeatMap(str_df.ic,
                x.field='occ.qtl',
                y.field='rate.qtl',
                pref.field='str.act.pref',
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))
  qtl.hm+ ylab('Nightly  (Pcntl)')

  qtl.hm.svm <- abbHeatMap(str_df.ic,
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T, 
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))
  qtl.hm.svm+ ylab('Nightly  (Pcntl)')

### Submarkets

Now we move on to looking more deeply at various submarkets.  We start by analyzing the differences between houses and apartments.

First we make a comparison table of the results.

 # Make Basic comparison table
  pref.table.type <- abbCreateCompTable(ic.df=str_df.ic,
                                   split.field='type')

 # Limit to the extrapolated pref type
  pref.table.type <- pref.table.type[pref.table.type$rev.type == 'Actual', ]

We then turn this into a simple chart to see the differences.

  ggplot(pref.table.type, aes(x=ID, weights=Var, fill=ID)) + 
    geom_bar() +
    scale_fill_manual(values=c('red', 'blue')) +
    xlab('') +
    ylab('% of properties where Airbnb is more profitable') +
    scale_y_continuous(breaks=c(0,.25,.5,.75,1),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none')

We then look at the relationship between occ rate and preference.  Before we start though we recalculate the percentiles within each submarket.

```{r build_occ_type1, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE}

 ## Create dataset

  type.data.ic <- str_df.ic

  # Id the types
  apt.id <- which(str_df.ic$type == 'Apartment')
  house.id <- which(str_df.ic$type == 'House')

  # Recalculate the percentiles
  type.data.ic$occ.qtl[apt.id] <- makeWtdQtl(type.data.ic$occ.rate[apt.id],
                                      return.type='rank')
  type.data.ic$occ.qtl[house.id] <- makeWtdQtl(
    type.data.ic$occ.rate[house.id], return.type='rank')

  type.data.ic$rate.qtl[apt.id] <- makeWtdQtl(type.data.ic$med.rate[apt.id],
                                  return.type='rank')
  type.data.ic$rate.qtl[house.id] <- makeWtdQtl(
    type.data.ic$med.rate[house.id], return.type='rank')

Then we calculate the preference figures

```{r build_occ_type, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE}

 ## Calc the occ rate numbers

  rawocc.type <- abbPrefPlot(type.data.ic,
                         x.field='occ.rate',
                         split.field='type',
                         smooth=TRUE,
                         smooth.span=.5 ,
                         spl.col=c('black', 'grey40'))

  pctocc.type <- abbPrefPlot(type.data.ic,
                         x.field='occ.qtl',
                         split.field='type',
                         smooth=TRUE,
                         smooth.span=.5 ,
                         spl.col=c('black', 'grey40'))

And then plot.

```{r plot_occ_type, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE, fig.width=7, fig.height=3.5, fig.cap='\\label{smocc}Preference vs Occ. Rates: By Type'}

 ## Build raw plot

  ptr.occ <- ggplot(rawocc.type$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate') +
    scale_color_manual(values=c('black', 'grey60'),
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Raw Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

 ## Build Percentile Plots

  ptq.occ <- ggplot(pctocc.type$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Pcntl)') +
    scale_color_manual(values=c('black', 'grey60'),
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Percentile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

 ## Plot

  ggMultiPlots(ptr.occ, ptq.occ, cols=2)

Finally, we make percentile SVM heatmaps for both apartments and houses.

```{r plot_svm_type, fig.width=8, fig.height=3.5, echo=FALSE, fig.cap='\\label{smhm}Analysis of Percentile Rates: By Type'}

 ## Make House SVM Heatmap

  house.hm.svm <- abbHeatMap(type.data.ic[type.data.ic$type ==
                                          'House', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5])) +
  theme(legend.position='none') +
  annotate("text", x = 50, y = 50, label = "Houses", size=9) +
  ylab('Nightly Rate (Pcntl)') +
  xlab('Occupancy Rate (Pcntl)')

 ## Make Apartment SVM Heat Map

  apt.hm.svm <- abbHeatMap(type.data.ic[type.data.ic$type ==
                                          'Apartment', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
  annotate("text", x = 50, y = 50, label = "Apartments", size=9)+
  ylab('Nightly Rate (Pcntl)') +
  xlab('Occupancy Rate (Pcntl)')

 ## Plot

  ggMultiPlots(house.hm.svm, apt.hm.svm, cols=2)

#### Geographic Submarkets

We now repeat the three analyses -- table, occupancy plot, heat map -- for the five geographic submarkets.

```{r table_geo, message=FALSE, warning=FALSE, results='asis', comment=FALSE, cache=TRUE, echo=FALSE, fig.cap='Geographic Submarket Preferability'}

 # Make Basic comparison table
  pref.table <- abbCreateCompTable(ic.df=str_df.ic,
                                   split.field='geo.mrkt')

 # Convert to wide format for output
  pref.table.wide <- dcast(pref.table, ID ~ rev.type, value.var='Var')

 # Limit to the extrapolated pref type
  pref.table <- pref.table[pref.table$rev.type == 'Actual', ]
  pref.table[ ,1] <- c('City-Core', 'City', 'Suburban', 'Rural', 'Beach')
  pref.table[ ,2] <- paste0(sprintf('%.1f', pref.table[ ,2] * 100), '%')

 # Fix names  
  names(pref.table) <- c('Sub-Market', 'Short-Term Preferable')
  options(xtable.comment = FALSE)

 ## Make printable table

  print(xtable(pref.table[,1:2], 
               caption='Geographic Submarket Preferability\n',
               digits=1,
               align='llr'),
        include.rownames=FALSE,
        caption.placement='top')

```{r build_occ_geo, fig.width=8, fig.height=4, echo=FALSE}

 ## Build data

  subm.data.ic <- str_df.ic

  # Get IDs
  cc.id <- which(str_df.ic$geo.mrkt == 'city-core')
  city.id <- which(str_df.ic$geo.mrkt == 'city')
  sub.id <- which(str_df.ic$geo.mrkt == 'suburban')
  rural.id <- which(str_df.ic$geo.mrkt == 'rural')
  beach.id <- which(str_df.ic$geo.mrkt == 'beach')

  # Recalc the percentiles
  subm.data.ic$occ.qtl[cc.id] <- makeWtdQtl(subm.data.ic$occ.rate[cc.id],
                                      return.type='rank')
  subm.data.ic$rate.qtl[cc.id] <- makeWtdQtl(subm.data.ic$med.rate[cc.id],
                                  return.type='rank')

  subm.data.ic$occ.qtl[city.id] <- makeWtdQtl(
      subm.data.ic$occ.rate[city.id], return.type='rank')
  subm.data.ic$rate.qtl[city.id] <- makeWtdQtl(
    subm.data.ic$med.rate[city.id], return.type='rank')

  subm.data.ic$occ.qtl[sub.id] <- makeWtdQtl(subm.data.ic$occ.rate[sub.id],
                                      return.type='rank')
  subm.data.ic$rate.qtl[sub.id] <- makeWtdQtl(subm.data.ic$med.rate[sub.id],
                                  return.type='rank')

  subm.data.ic$occ.qtl[rural.id] <- makeWtdQtl(subm.data.ic$occ.rate[rural.id],
                                      return.type='rank')
  subm.data.ic$rate.qtl[rural.id] <- makeWtdQtl(subm.data.ic$med.rate[rural.id],
                                  return.type='rank')

  subm.data.ic$occ.qtl[beach.id] <- makeWtdQtl(subm.data.ic$occ.rate[beach.id],
                                      return.type='rank')
  subm.data.ic$rate.qtl[beach.id] <- makeWtdQtl(subm.data.ic$med.rate[beach.id],
                                  return.type='rank')

```{r plot_occ_geo, fig.width=9, fig.height=4, message=FALSE, warning=FALSE, echo=FALSE, fig.cap='\\label{smocc}Analysis of Percentile Rates: Geographic Submarkets'}

 ## Set colors

  sm.col <- abb.col[c(1, 3, 6, 5, 2)]

 ## Create plot data

  rawocc.geo <- abbPrefPlot(subm.data.ic,
                        x.field='occ.rate',
                        split.field='geo.mrkt',
                        smooth=TRUE,
                        smooth.span=.5 ,
                        spl.col=sm.col)

  pctocc.geo <- abbPrefPlot(subm.data.ic,
                         x.field='occ.qtl',
                         split.field='geo.mrkt',
                         smooth=TRUE,
                         smooth.span=.5 ,
                         spl.col=sm.col)

 ## Make plot objects

  smr.occ <- ggplot(rawocc.geo$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate') +
    scale_color_manual(values=sm.col,
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Raw Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

  smq.occ <- ggplot(pctocc.geo$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Pcntl)') +
    scale_color_manual(values=sm.col,
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Percentile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

 ## Plot

  ggMultiPlots(smr.occ, smq.occ, cols=2)

```{r build_svm_geo, fig.width=8, fig.height=4, echo=FALSE, fig.cap='Analysis of Percentile Rates: Metro Level'}

  # Build the SVM Heatmaps

  core.hm.svm <- abbHeatMap(subm.data.ic[subm.data.ic$geo.mrkt ==
                                          'city-core', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5])) +
  theme(legend.position='none') +
    ylab('') +
    xlab('') +
  annotate("text", x = 50, y = 50, label = "City-Core", size=9)

  city.hm.svm <- abbHeatMap(subm.data.ic[subm.data.ic$geo.mrkt ==
                                          'city', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
    ylab('') +
    xlab('')+
  annotate("text", x = 50, y = 50, label = "City", size=9)

  sub.hm.svm <- abbHeatMap(subm.data.ic[subm.data.ic$geo.mrkt ==
                                          'suburban', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
    ylab('') +
    xlab('')+
  annotate("text", x = 50, y = 50, label = "Suburban", size=9)

  rural.hm.svm <- abbHeatMap(subm.data.ic[subm.data.ic$geo.mrkt ==
                                          'rural', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
    ylab('') +
    xlab('')+
  annotate("text", x = 50, y = 50, label = "Rural", size=9)

  beach.hm.svm <- abbHeatMap(subm.data.ic[subm.data.ic$geo.mrkt ==
                                          'beach', ],
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.act.pref',
                   svm=T,
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))+
  theme(legend.position='none')+
    ylab('') +
    xlab('')+
  annotate("text", x = 50, y = 50, label = "Beach", size=9)

```{r plot_svm_geo, message=FALSE, warning=FALSE, fig.width=8, fig.height=8, echo=FALSE, fig.cap='\\label{smhmps}Geographic Submarket Heatmaps'}

  ggMultiPlots(core.hm.svm, city.hm.svm, sub.hm.svm,
               rural.hm.svm, beach.hm.svm, cols=2)

@@

  str_df.ic$bedbath <- factor(str_df.ic$bedbath,
                                levels=c('1..1', '2..1', '2..2',
                                         '3..1', '3..2', '4..2'))

  mod.act.str <- glm(str.act.pref~type+as.factor(bedbath),
                     family=binomial(link='logit'),
                     data=str_df.ic)
  mod.pot.str <- glm(str.pot.pref~type+as.factor(bedbath),
                     family=binomial(link='logit'),
                     data=str_df.ic)
```

@@

  mod.act.loc <- glm(str.act.pref~type+as.factor(bedbath)+
                                  geo.mrkt,
                     family=binomial(link='logit'),
                     data=str_df.ic)
  mod.pot.loc <- glm(str.pot.pref~type+as.factor(bedbath)+
                                  geo.mrkt,
                     family=binomial(link='logit'),
                     data=str_df.ic)
```

@@

  mod.act.host <- glm(str.act.pref~type+as.factor(bedbath)+
                        geo.mrkt+
                        I(max.guests/bedrooms)+min.stay+
                        I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                      family=binomial(link='logit'),
                      data=str_df.ic)
  mod.pot.host <- glm(str.pot.pref~type+as.factor(bedbath)+
                        geo.mrkt+
                        I(max.guests/bedrooms)+min.stay+
                        I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                      family=binomial(link='logit'),
                      data=str_df.ic)
```

We combine all model results in to a list for easy export.

  pref.models <- list(mod.act.str, mod.act.loc, mod.act.host, 
                      mod.pot.str, mod.pot.loc, mod.pot.host)
  names(pref.models) <- c('act.str', 'act.loc', 'act.host',
                          'pot.str', 'pot.loc', 'pot.host')

We then also run a set of regression models where the premium (or discount) for short-term leasing is the dependent variable.  We use the same three model specifications as above.

  mod.act.str.prem <- lm(str.act.prem~type+as.factor(bedbath),
                    data=str_df.ic)
  mod.pot.str.prem <- lm(str.pot.prem~type+as.factor(bedbath),
                    data=str_df.ic)

  mod.act.loc.prem <- glm(str.act.pref~type+as.factor(bedbath)+
                                  geo.mrkt,
                     family=binomial(link='logit'),
                     data=str_df.ic)
  mod.pot.loc.prem <- glm(str.pot.pref~type+as.factor(bedbath)+
                                  geo.mrkt,
                     family=binomial(link='logit'),
                     data=str_df.ic)

  mod.act.host.prem <- lm(str.act.prem~type+as.factor(bedbath)+
                        geo.mrkt+
                        I(max.guests/bedrooms)+min.stay+
                        I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                      data=str_df.ic)
  mod.pot.host.prem <- lm(str.pot.prem~type+as.factor(bedbath)+
                        geo.mrkt+
                        I(max.guests/bedrooms)+min.stay+
                        I(cancel.policy=='Flexible') + I(cancel.policy=='Strict'),
                      data=str_df.ic)
```

We then package these into a list for easy export

    prem.models <- list(mod.act.str.prem, mod.act.loc.prem, mod.act.host.prem, 
                      mod.pot.str.prem, mod.pot.loc.prem, mod.pot.host.prem)
    names(prem.models) <- c('act.str', 'act.loc', 'act.host',
                          'pot.str', 'pot.loc', 'pot.host')

## Sensitivity

### Potential Revenue

To test for the sensitivity of the result to the revenue type, we re-do some of the intial data vizs.  First we build a new preference table looking at the submarkets of type and geography.

```{r pot_table, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='Potential Short-Term Preference'}

 # Make Basic comparison table of type
  tpref.table <- abbCreateCompTable(ic.df=str_df.ic,
                                    split.field='type')
 # Extract Potential Results
  tpref.table <- tpref.table[tpref.table$rev.type == 'Potential', ]

 # Make baseic comparison table of submarkets  
  spref.table <- abbCreateCompTable(ic.df=str_df.ic,
                                    split.field='geo.mrkt')

 # Extract potential results  
  spref.table <- spref.table[spref.table$rev.type == 'Potential', ]

 # Combine and round  
  st.table <- rbind(tpref.table, spref.table)
  st.table[,2] <- paste0(sprintf('%.1f', st.table[,2]*100), '%')

 ## Fix names and print

  names(st.table) <- c('Sub-Market', 'Short-Term Preferable')
  options(xtable.comment = FALSE)
  print(xtable(st.table[,1:2], 
               caption='Potential Revenue Preferability\n',
               digits=1, align='llr'),
               include.rownames=FALSE,
               caption.placement='top',
               hline.after=c(-1, 0, 2, 7, 7))

Next we recrate the SVM percentile heat maps.

```{r pot_svm, results='asis', echo=FALSE, message=FALSE, comment=FALSE}

 ## create data and re-estimate the percentiles

  pot.data.ic <- str_df.ic

  pot.data.ic$occ.qtl <- makeWtdQtl(pot.data.ic$pot.occ.rate, return.type='rank')

 ## Make the pecentile heatmap

  pqtl.hm <- abbHeatMap(pot.data.ic,
                x.field='occ.qtl',
                y.field='rate.qtl',
                pref.field='str.pot.pref',
                svm=F, 
                alpha.count=T,
                add.points=T,
                fill.colors=c(abb.col[1], abb.col[5]))
  pqtl.hm <- pqtl.hm + ylab('Nightly Rate (Qtl)')

 ## Make percentile SVM heatmap

  pqtl.hm.svm <- abbHeatMap(pot.data.ic,
                   x.field='occ.qtl',
                   y.field='rate.qtl',
                   pref.field='str.pot.pref',
                   svm=T, 
                   alpha.count=F,
                   add.points=T,
                   quantile=T,
                   fill.colors=c(abb.col[1], abb.col[5]))
  pqtl.hm.svm <- pqtl.hm.svm + ylab('Nightly Rate (Pcntl)')+
    xlab('Occupancy Rate (Pcntl)') +
  annotate("text", x = 50, y = 50, label = "Potential", size=9)

 ## Add annotations to previous SVM

>>>>

---
title: "Should I Airbnb My Property: Analytic Code Explanation"
author: "Andy Krause and Gideon Aschwanden"
date: "March 10, 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

## Load Data

```{r load_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  load(file.path(data_path, 'prepared', 'prepared_data.RData'))

@@

  # Set cost per person per day
  cpppd <- 4.5

  # Calculate avg number of guests
  str_tdf$avg_guests <- str_tdf$bedrooms * 1.5

@@

  str_tdf$act_revenue <- ((str_tdf$bookings * str_tdf$med_rate) -
                                 (str_tdf$bookings * str_tdf$avg_guests * cpppd))

```{r lik_revenues, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Create Extrapolation parameter
  str_tdf$extr_par <- 366 / str_tdf$total_days

  # Calculate likely revenue
  str_tdf$lik_revenue <- ((str_tdf$bookings * str_tdf$extr_par * str_tdf$med_rate) -
                           (str_tdf$bookings * str_tdf$avg_guests *
                             cpppd * str_tdf$extr_par))
```

@@

  str_tdf$occ_rate <- 1 - (str_tdf$block_rate + str_tdf$avail_rate)

  str_tdf$pot_occ_rate <- ((str_tdf$bookings * str_tdf$extr_par) /
                             (366 - (366 * str_tdf$block_rate)))

  str_tdf$pot_revenue <- ((str_tdf$pot_occ_rate * 366 * str_tdf$med_rate) -
                           (str_tdf$pot_occ_rate * 366 * str_tdf$avg_guests * cpppd))

@@

```{r flatten_ltr}

  ltr_df <- ltr_tdf %>% 
    unnest() %>%
    select(-suburb.x) %>%
    rename(suburb=suburb.y)

@@

  ltr_mod_spec <- formula(log(price) ~ as.factor(type) +
                            as.factor(bedbath) + as.factor(suburb) +
                            as.factor(month))

@@

  str_tdf$tbg <- paste0(str_tdf$product,'_', str_tdf$geo_mrkt)
  ltr_df$tbg <- paste0(ltr_df$product,'_', ltr_df$geo_mrkt)

  imp_df <- abbImputeCompare(str.df=str_tdf, 
                                ltr.df=ltr_df, 
                                mod.spec=ltr_mod_spec, 
                                match.factor='suburb',
                                split.field='tbg', 
                                verbose=FALSE)

@@

  # Extract out DFs (only those that imputed)
  str_tdf <- str_tdf %>%
    dplyr::left_join(imp_df, by='property_id') %>%
    dplyr::mutate(act_pref = ifelse(act_revenue > ltr_imp_revenue, 1, 0),
                  lik_pref = ifelse(lik_revenue > ltr_imp_revenue, 1, 0),
                  pot_pref = ifelse(pot_revenue > ltr_imp_revenue, 1, 0))

@@

  glm_lik_str <- glm(lik_pref ~ type + as.factor(bedbath),
                     family=binomial(link='logit'),
                     data=str_tdf)

  glm_pot_str <- glm(pot_pref~type+as.factor(bedbath),
                       family=binomial(link='logit'),
                       data=str_tdf)
```

@@

  glm_lik_loc <- glm(lik_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)

  glm_pot_loc <- glm(pot_pref ~ type + as.factor(bedbath)+ as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)
```

@@

  glm_lik_host <- glm(lik_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + as.factor(canc_policy),
                      family=binomial(link='logit'),
                      data=str_tdf)

  glm_pot_host <- glm(pot_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + as.factor(canc_policy),
                      family=binomial(link='logit'),
                      data=str_tdf)

  lik_glms <- list(str=glm_lik_str,
                   loc=glm_lik_loc,
                   host=glm_lik_host)

  pot_glms <- list(str=glm_pot_str,
                   loc=glm_pot_loc,
                   host=glm_pot_host)
